---
title: "MLD"
author: "Ankita Pathak, Nancy Jain and Sanyukta Ghai"
date: "02/26/2019"
output:
  word_document: default
  html_document: default
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
---

Research Questions
=====================================================

Can we conclude that mortgage lending institutions discriminate against minorities?  


```{r echo = FALSE}
# Course: ECON 
# Title: MLD_Project
# Purpose: EDA
# Date: February 26 , 2019
# Author: Ankita Pathak, Nancy Jain and Sanyukta Ghai
```

```{r echo = FALSE, warning = FALSE, message=FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

```{r echo = FALSE, warning = FALSE, message=FALSE}

library(tidyverse)
library(here)
library(kableExtra)
library(knitr)
library(stringr)
library(psych)

knitr::opts_chunk$set(echo = FALSE)
library(stargazer)

```

Load data and begin 
=========================================
Work with `MLD Data File-1.csv` data set

```{r, echo = FALSE}
# Load data 
mld_data <- read.csv(here("raw_data", "MLD Data File-1.csv"))

#colnames(mld_data)

#head(mld_data, 10)

# Structure of data so know what working with
#str(mld_data)

#summary(mld_data)

# Married .?
# Male .?
# Guidelines 666?
# LOANPRC - Loan to Value Multiply by 100

#The data set includes the following variables:
#APPROVE = 1 if mortgage loan was approved, = 0 otherwise
#GDLIN = 1 if credit history meets guidelines, = 0 otherwise
#LOANPRC = loan amount/purchase price
#OBRAT = other obligations as a percent of total income
#MALE = 1 if male, = 0 otherwise
#MARRIED = 1 if married, = 0 otherwise
#BLACK = 1 if black, = 0 otherwise
#HISPAN = 1 if Hispanic, = 0 otherwise
#(All applicants are non-Hispanic white, non Hispanic black, or Hispanic.  In 1990 about 94% of Boston residents were white, #Black, or Hispanic. (Links to an external site.)Links to an external site.)

```


Filtering required data
=======================

```{r}

options(scipen = 999)

base <- mld_data %>%
  filter(GDLIN == 1 | GDLIN == 0,
         LOANPRC <= 1) %>%
  mutate(LOANPRC = 100 * LOANPRC,
         GDLIN = if_else(GDLIN == 1, TRUE, FALSE),
         WHITE = if_else(BLACK == 0 & HISPAN == 0, TRUE, FALSE),
         BLACK = if_else(BLACK == 1, TRUE, FALSE),
         HISPAN = if_else(HISPAN == 1, TRUE, FALSE),
         APPROVE = if_else(APPROVE == 1, TRUE, FALSE)) %>%
  select(APPROVE, GDLIN, OBRAT, LOANPRC, WHITE, BLACK, HISPAN)

#head(base)

```

### Sample Selection Criteria

+ GDLIN should be between 0 or 1(meaning either a loan requester meets the guidelines or does not) 
  Removed arbitrary values like 666  
+ LOANPRC less than 100%. Why would the bank approve loans greater than 100% loan to value?  


Subset all race and ethnicities
===============================

```{r}

#str(base)

summary(base)

base_white <- base %>%
  filter(WHITE == TRUE)

base_black <- base %>%
  filter(BLACK == TRUE)

base_hispanic <- base %>%
  filter(HISPAN == TRUE)

```

## Descriptive Statistics for Entire Sample


```{r}
stargazer(as.data.frame(base), align = FALSE, 
          title = "Descriptive Statistics for Entire sample", 
          omit.summary.stat = c("n","p25","p75"), 
          notes = "Total number of records - 1948",
          covariate.labels=c("Mortgage loan approval",
                             "Credit history meets guidelines", 
                             "Other obligations as a percent of total income", 
                             "Loan amount as a percent of purchase price",
                             "Non-Hispanic White",
                             "Non-Hispanic Black",
                             "Hispanic",
                             "Intercept"),
          header = FALSE,
          summary.logical = TRUE,
          type = "text")

```

```{r}

desc_stats_sample <- base %>%
  mutate_if(is.logical,as.numeric) %>%
  describe(interp=FALSE,skew = TRUE, ranges = TRUE,trim=.1,
           type=3,check=TRUE,fast=NULL,quant=NULL,IQR=FALSE,omit=FALSE) #%>%
  
  #select(mean, sd, median, min, max) %>%
  #mutate_all(funs(round(.,2)))
desc_stats_sample <- data.frame(Variables = row.names(desc_stats_sample), desc_stats_sample) %>%
  select(Variables, mean, sd, median, min, max) %>%
  mutate_if(is.numeric, round, 2)

desc_stats_sample <- desc_stats_sample %>%
  mutate(Variables = c("Mortgage Loan Approval",
                    "Credit history meets guidelines", 
                    "Other obligations as a percent of total income", 
                    "Loan amount as a percent of purchase price",
                    "Non-Hispanic White",
                    "Non-Hispanic Black",
                    "Hispanic"))

desc_stats_sample %>%
  kable("latex", booktabs = T, caption = "Descriptive Statistics for Entire sample") %>%
  kable_styling(position = "center", full_width = F) %>%
  column_spec(1, width = "8cm") %>%
  footnote(general = c("Total number of records: 1954"))   

```



```{r}

desc_stats_white <- base_white %>%
  mutate_if(is.logical,as.numeric) %>%
  describe(interp=FALSE,skew = TRUE, ranges = TRUE,trim=.1,
           type=3,check=TRUE,fast=NULL,quant=NULL,IQR=FALSE,omit=FALSE) #%>%

#desc_stats_white

desc_stats_white <- data.frame(Variables = row.names(desc_stats_white), desc_stats_white) %>%
  select(Variables, mean, sd, median, min, max) %>%
  mutate_if(is.numeric, round, 2)

desc_stats_white <- desc_stats_white %>%
  mutate(Variables = c("Mortgage Loan Approval",
                    "Credit history meets guidelines", 
                    "Other obligations as a percent of total income", 
                    "Loan amount as a percent of purchase price",
                    "Non-Hispanic White",
                    "Non-Hispanic Black",
                    "Hispanic"))

desc_stats_white %>%
  kable("latex", booktabs = T, caption = "Descriptive Statistics for Non-Hispanic White sample") %>%
  kable_styling(position = "center", full_width = F) %>%
  column_spec(1, width = "8cm") %>%
  footnote(general = c("Total number of records: 1653"))

```

```{r}
desc_stats_black <- base_black %>%
  mutate_if(is.logical,as.numeric) %>%
  describe(interp=FALSE,skew = TRUE, ranges = TRUE,trim=.1,
           type=3,check=TRUE,fast=NULL,quant=NULL,IQR=FALSE,omit=FALSE) #%>%

#desc_stats_black

desc_stats_black <- data.frame(Variables = row.names(desc_stats_black), desc_stats_black) %>%
  select(Variables, mean, sd, median, min, max) %>%
  mutate_if(is.numeric, round, 2)

desc_stats_black <- desc_stats_black %>%
  mutate(Variables = c("Mortgage Loan Approval",
                    "Credit history meets guidelines", 
                    "Other obligations as a percent of total income", 
                    "Loan amount as a percent of purchase price",
                    "Non-Hispanic White",
                    "Non-Hispanic Black",
                    "Hispanic"))

desc_stats_black %>%
  kable("latex", booktabs = T, caption = "Descriptive Statistics for Non-Hispanic Black sample") %>%
  kable_styling(position = "center", full_width = F) %>%
  column_spec(1, width = "8cm") %>%
  footnote(general = c("Total number of records: 194"))
```

```{r}


desc_stats_hispanic <- base_hispanic %>%
  mutate_if(is.logical,as.numeric) %>%
  describe(interp=FALSE,skew = TRUE, ranges = TRUE,trim=.1,
           type=3,check=TRUE,fast=NULL,quant=NULL,IQR=FALSE,omit=FALSE) #%>%

#desc_stats_hispanic

desc_stats_hispanic <- data.frame(Variables = row.names(desc_stats_hispanic), desc_stats_hispanic) %>%
  select(Variables, mean, sd, median, min, max) %>%
  mutate_if(is.numeric, round, 2)

desc_stats_hispanic <- desc_stats_hispanic %>%
  mutate(Variables = c("Mortgage Loan Approval",
                    "Credit history meets guidelines", 
                    "Other obligations as a percent of total income", 
                    "Loan amount as a percent of purchase price",
                    "Non-Hispanic White",
                    "Non-Hispanic Black",
                    "Hispanic"))

desc_stats_hispanic %>%
  kable("latex", booktabs = T, caption = "Descriptive Statistics for Hispanic sample") %>%
  kable_styling(position = "center", full_width = F) %>%
  column_spec(1, width = "8cm") %>%
  footnote(general = c("Total number of records: 107"))

```


## Observations

+ On average, a mortgage loan approval rate is 87.9% with 91.4% of applications meeting credit history guidelines
+ On average, people have obligations of \$33, if their total income is $100
+ On average, people are applying for 76% loan amount, where 76% is calculated as amount applied for, as a percent of         value of property
+ Highest percentage of people applying for a loan are non-hispanic whites with 84.6%
+ Non-hispanic black with 10% and hispanic with 5.4% people applying for a loan

## Descriptive Statistics for White Sample


```{r}
stargazer(as.data.frame(base_white), align = FALSE, 
          title = "Descriptive Statistics for WHITE sample", 
          omit.summary.stat = c("n","p25","p75"), 
          notes = "Total number of Non-Hispanic Whites - 1648",
          covariate.labels=c("Mortgage loan approval",
                             "Credit history meets guidelines", 
                             "Other obligations as a percent of total income", 
                             "Loan amount as a percent of purchase price",
                             "Non-Hispanic White",
                             "Non-Hispanic Black",
                             "Hispanic",
                             "Intercept"),
          header = FALSE,
          summary.logical = TRUE,
          type = "text")
```

## Observations

+ On average, a mortgage loan approval rate for non-hispanic whites is 91%
+ 94% of applications meet credit history guidelines
+ On average, non-hispanic whites have obligations of \$32, if their total income is $100
+ On average, non-hispanic whites are applying for 74% loan amount, where 74% is calculated as amount applied for, as a        percent of value of property

## Descriptive Statistics for Black Sample


```{r}
stargazer(as.data.frame(base_black), align = FALSE, 
          title = "Descriptive Statistics for BLACK sample", 
          omit.summary.stat = c("n","p25","p75"), 
          notes = "Total number of Non-Hispanic Blacks - 194",
          covariate.labels=c("Mortgage loan approval",
                             "Credit history meets guidelines", 
                             "Other obligations as a percent of total income", 
                             "Loan amount as a percent of purchase price",
                             "Non-Hispanic White",
                             "Non-Hispanic Black",
                             "Hispanic",
                             "Intercept"),
          header = FALSE,
          summary.logical = TRUE,
          type = "text")
```

## Observations

+ On average, a mortgage loan approval rate for non-hispanic blacks is 67%
+ Only 73% of applications meet credit history guidelines
+ On average, non-hispanic blacks have obligations of \$35, if their total income is $100
+ On average, non-hispanic blacks are applying for 83% loan amount, where 83% is calculated as amount applied for, as a        percent of value of property

## Descriptive Statistics for Hispanic Sample


```{r}
stargazer(as.data.frame(base_hispanic), align = FALSE, 
          title = "Descriptive Statistics for HISPANIC sample",
          omit.summary.stat = c("n","p25","p75"), 
          notes = "Total number of Hispanics - 106",
          covariate.labels=c("Mortgage loan approval",
                             "Credit history meets guidelines", 
                             "Other obligations as a percent of total income", 
                             "Loan amount as a percent of purchase price",
                             "Non-Hispanic White",
                             "Non-Hispanic Black",
                             "Hispanic",
                             "Intercept"),
          header = FALSE,
          summary.logical = TRUE,
          type = "text")
```

## Observations

+ On average, a mortgage loan approval rate for hispanics is 78.3%
+ Approximately, 87% of applications meet credit history guidelines
+ On average, hispanics have obligations of \$33, if their total income is $100
+ On average, hispanics are applying for 83% loan amount, where 83% is calculated as amount applied for, as a        percent of value of property

### Descriptive Observations

+ There are 1648 Non-Hispanic White  
+ There are 194 Non-Hispanic Black  
+ There are 106 Hispanics.

## Probit Model

```{r}

Approval_probit <- glm(APPROVE ~ GDLIN + OBRAT + LOANPRC + BLACK + HISPAN, family=binomial(link="probit"), data=base) 

summary(Approval_probit)

```

## Results of Probit Models

```{r warning=FALSE}
stargazer(Approval_probit, 
          title="Results", 
          align=TRUE, 
          type = "text",
          column.labels=c("Probit"),
          dep.var.labels=c("Mortgage loan approval"),
          covariate.labels=c("Credit history meets guidelines", 
                             "Other obligations as a percent of total income", 
                             "Loan amount as a percent of purchase price",
                             "Non-Hispanic Black",
                             "Hispanic",
                             "Intercept"),
          notes.append = TRUE, notes.align = "l",
          notes = c("Reference Category for Race is a Non-Hispanic White"), 
          header = FALSE)

```


```{r}

Approval_probit_table <- data.frame(Variables = row.names(summary(Approval_probit)$coeff), summary(Approval_probit)$coeff)



colnames(Approval_probit_table) <- c('Variables', 'Estimates', "Std. Error", "z value", "p_value")

#Approval_probit_table
Approval_probit_table <- Approval_probit_table %>%
  mutate(p = if_else(p_value < 0.01, "***", if_else(p_value < 0.05, "**", if_else(p_value < 0.10, "*", "")))) %>%
  mutate(Variables = c("Intercept",
                    "Credit history meets guidelines", 
                    "Other obligations as a percent of total income", 
                    "Loan amount as a percent of purchase price",
                    "Non-Hispanic Black",
                    "Hispanic")) %>%
    mutate(Estimates = round(Estimates, 3),
         `Std. Error` = round(`Std. Error`, 3),
         `z value` = round(`z value`, 3),
         p_value = round(p_value, 3)) %>%
  mutate(p = if_else(p_value < 0.01, "***", if_else(p_value < 0.05, "**", if_else(p_value < 0.10, "*", "")))) %>%
  select(Variables, Estimates, `Std. Error`, `z value`, `p value` = p_value, " " = p )  


Approval_probit_table %>%
  kable("latex", booktabs = T, caption = "Probit Model") %>%
  kable_styling(position = "center", full_width = F) %>%
  column_spec(1, width = "8cm") %>%
  add_header_above(c(" " = 1, "Dependent Variable : \nMortgage loan approval" = 5)) %>%
  footnote(general = c("*p<0.1; **p<0.05; ***p<0.01", 
                       "Reference Category for Race is a Non-Hispanic White", 
                       "Total number of observations: 1954",
                       "Log Likelihood: -467.078"))
```


## Observations

1. Non-Hispanic Black: The coefficient of Non-Hispanic Black is -0.449. A person having origins in any of the black racial groups like non-hispanic black, versus a person who is non-hispanic white (the reference group), decreases the z-score by 0.449 

2. Hispanic: The coefficient of Hispanic is -0.413. A person having origins in any of the hispanic groups, versus a person who is a non-hispanic (the reference group), decreases the z-score by 0.413


## Logit Model

```{r}

Approval_logit <- glm(APPROVE ~ GDLIN + OBRAT + LOANPRC + BLACK + HISPAN, family=binomial(link="logit"), data=base)

summary(Approval_logit)

Approval_logit_table <- data.frame(Variables = row.names(summary(Approval_logit)$coeff), summary(Approval_logit)$coeff)

colnames(Approval_logit_table) <- c('Variables', 'Estimates', "Std. Error", "z value", "p_value")

Approval_logit_table <- Approval_logit_table %>%
  mutate(p = if_else(p_value < 0.01, "***", if_else(p_value < 0.05, "**", if_else(p_value < 0.10, "*", ""))))

Approval_logit.OR <-  exp(coef(Approval_logit)) %>%
  as.data.frame()

Approval_logit.OR <- data.frame(Variables = row.names(Approval_logit.OR), Approval_logit.OR)

colnames(Approval_logit.OR) <- c('Variables', 'Odds Ratio')
#Approval_logit.OR

Approval_logit_table <- Approval_logit_table %>%
  left_join(Approval_logit.OR) %>%
  mutate(Estimates = round(Estimates, 3),
         `Odds Ratio` = round(`Odds Ratio`, 3),
         `Std. Error` = round(`Std. Error`, 3),
         `z value` = round(`z value`, 3),
         #p_value = str_c(round(p_value, 3), p)) %>%
         p_value = round(p_value, 3)) %>%
  #select(-(p))
  #select(Variables, Estimates, `Odds Ratio`, `Std. Error`, `z value`, `p value` = p_value )
  select(Variables, Estimates, `Std. Error`, `Odds Ratio`, `p value` = p_value, " " = p )

Approval_logit_table <- Approval_logit_table %>%
  mutate(Variables = c("Intercept",
                    "Credit history meets guidelines", 
                    "Other obligations as a percent of total income", 
                    "Loan amount as a percent of purchase price",
                    "Non-Hispanic Black",
                    "Hispanic"))
Approval_logit_table
```


## Results of Logit using Kable Extra

```{r}
Approval_logit_table %>%
  kable("latex", booktabs = T, caption = "Logit Model") %>%
  kable_styling(position = "center", full_width = F) %>%
  column_spec(1, width = "8cm") %>%
  add_header_above(c(" " = 1, "Dependent Variable : \nMortgage loan approval" = 4)) %>%
  footnote(general = c("*p<0.1; **p<0.05; ***p<0.01", 
                       "Reference Category for Race is a Non-Hispanic White", 
                       "Total number of observations: 1954"))
#%>%
 # kable_as_image()

```



## Results of Logit Models

```{r warning=FALSE, include=FALSE}
stargazer(Approval_logit,
          title="Results", 
          align=TRUE, 
          type = "text",
          column.labels=c("Estimates"),
          dep.var.labels=c("Mortgage loan approval"),
          covariate.labels=c("Credit history meets guidelines", 
                             "Other obligations as a percent of total income", 
                             "Loan amount as a percent of purchase price",
                             "Non-Hispanic Black",
                             "Hispanic",
                             "Intercept"),
          notes.append = TRUE, notes.align = "l",
          notes = c("Reference Category for Race is a Non-Hispanic White"), 
          header = FALSE)

```

## Observations

## Median Values

```{r}

median_values <- base %>%
  summarise(median_OBRAT = median(OBRAT), medianLOANPRC = median(LOANPRC))

median_OBRAT <- median_values[1,1]

median_LOANPRC <- median_values[1,2]

```


## Prototypical Individuals

```{r}

get_prototypical_individual_df <- function(is_probit = FALSE, coeff_list){

  # Extract Coefficient Constants
  coeff_Intercept <- coeff_list[1]
  coeff_GDLIN <- coeff_list[2]
  coeff_OBRAT <- coeff_list[3]
  coeff_LOANPRC <- coeff_list[4]
  coeff_NH_BLACK <- coeff_list[5]
  coeff_HISPANIC <- coeff_list[6]

  # Create an empty tibble
  df <- tibble(Meets_guidelines = integer(), NH_White = integer(), NH_Black = integer(), Hispanic = integer())

  # Add prototypical Individuals
  df <- df %>%
  add_row(Meets_guidelines = 1, NH_White = 1, NH_Black = 0, Hispanic = 0 ) %>%
  add_row(Meets_guidelines = 1, NH_White = 0, NH_Black = 1, Hispanic = 0 ) %>%
  add_row(Meets_guidelines = 1, NH_White = 0, NH_Black = 0, Hispanic = 1 ) %>%
  add_row(Meets_guidelines = 0, NH_White = 1, NH_Black = 0, Hispanic = 0 ) %>%
  add_row(Meets_guidelines = 0, NH_White = 0, NH_Black = 1, Hispanic = 0 ) %>%
  add_row(Meets_guidelines = 0, NH_White = 0, NH_Black = 0, Hispanic = 1 )
  
  # Add columns to define prototypical individuals in words
  df <- df %>%
    mutate(Prototype_GDLIN = if_else(Meets_guidelines == 0, "Does not meet guidelines", "Meets Guidelines"),
           Prototype_Race = str_c(if_else(NH_White == 0, "", "Non-Hispanic White"),
                           if_else(NH_Black == 0, "", "Non-Hispanic Black"),
                           if_else(Hispanic == 0, "", "Hispanic")))
  
  # Calculate z - differently for logit and probit based on value of is_probit
  df <- df %>%
    mutate(z = coeff_Intercept + 
             coeff_GDLIN * Meets_guidelines +
             coeff_OBRAT * median_OBRAT +
             coeff_LOANPRC * median_LOANPRC +
             coeff_NH_BLACK * NH_Black +
             coeff_HISPANIC * Hispanic)
  
  if (is_probit){
    df <- df %>%
      mutate(Approval_z_score = z)  %>%
      mutate(Predicted_prob_approval = round(pnorm(Approval_z_score),4))
  }
  else
  {
    df <- df %>%
      mutate(Approval_z_score = (1/(1+exp(-1*z)))) %>%
      mutate(Predicted_prob_approval = round(Approval_z_score,4))
  }
  
  # Calculate predicted probability of approval and select relevant columns
  df <- df  %>%
    select(Prototype_GDLIN, Prototype_Race, Predicted_prob_approval) 

  return(df)
}


```

```{r}
logit_prototypes <- get_prototypical_individual_df(FALSE, summary(Approval_logit)$coeff)

probit_prototypes <- get_prototypical_individual_df(TRUE, summary(Approval_probit)$coeff)

prototypes <- probit_prototypes %>%
  left_join(logit_prototypes, by = c("Prototype_GDLIN", "Prototype_Race")) %>%
  mutate(Probit = Predicted_prob_approval.x,
         Logit = Predicted_prob_approval.y
         ) %>%
  select(-c(Predicted_prob_approval.x, Predicted_prob_approval.y))

prototypes <- prototypes %>%
  select(`Credit Guidelines` = Prototype_GDLIN, Race = Prototype_Race, Probit, Logit )

prototypes %>%
  kable("latex", booktabs = T, caption = "Predicted Probability of Approval for Prototypic Individuals") %>%
  kable_styling(position = "center", full_width = F) %>%
  add_header_above(c(" " = 2, "Predicted Probability of Approval" = 2)) %>%
  column_spec(3, width = "2cm") %>%
  column_spec(4, width = "2cm") %>%
  collapse_rows(1, latex_hline = "full", valign = "middle") %>%
  footnote(general = c("Median Value of OBRAT : 33%", "Median Value of LOANPRC : 80%")) 
  
#%>%
 # group_rows("Meets Guidelines", 1, 3) %>%
  #group_rows("Does Not Meet Guidelines", 4, 6)

```

