---
title: "Zillow Analysis by Data-Cated"
author: "Eirik, Jillian, Nancy, Yifan, Ziyu "
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 3
    toc_float: yes
  # pdf_document: 
  #   latex_engine: xelatex
  
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	fig.width = 8.5,
	message = FALSE,
	warning = FALSE,
	# include = FALSE,
	echo = FALSE
)

options(tinytex.verbose = TRUE)

# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environment of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE) {
  lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), 
         detach, character.only = TRUE, unload = TRUE)
}
```


# Install and load Packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)

library(skimr)
library(sqldf)
```


# Load Data

```{r, warning=FALSE, message=FALSE, echo=TRUE}

agent_acq_dates <- read_csv(here("RawData", "AgentAcquisitionDates.csv"))
agent_lookup <- read_csv(here("RawData", "AgentIDZUIDLookup.csv"))
agent_leads <- read_csv(here("RawData", "AgentLeads.csv"))
phonecalls_16_17 <- read_csv(here("RawData", "PhoneCalls20161001_20170228.csv"))
phonecalls_17 <- read_csv(here("RawData", "PhoneCalls20170301_20170630.csv"))
sales_meeting <- read_csv(here("RawData", "SalesMeetings.csv"))

```


# Getting an overview of the data

```{r, warning=FALSE, message=FALSE, echo=TRUE}

summary(agent_acq_dates)
summary(agent_leads)
summary(phonecalls_16_17)
summary(phonecalls_17)
summary(sales_meeting)
skim(agent_acq_dates)


```


# Begin making on table with all the important variables - join agent_leads and agent_lookup

```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid <- agent_leads %>%
  left_join(agent_lookup, c("REAgentID" = "REAgentID"))

```


# Duplicate zuids?

```{r, warning=FALSE, message=FALSE, echo=TRUE}
n_occur <- data.frame(table(leads_zuid$REAgentID))
n_occur[n_occur$Freq > 50,]

```


```{r, warning=FALSE, message=FALSE, echo=TRUE}
n_occur <- data.frame(table(agent_lookup$ZUID))
n_occur[n_occur$Freq > 1,]

```


# Only select rows with acquisition date after 2016-09-30

```{r, warning=FALSE, message=FALSE, echo=TRUE}
acq <- agent_acq_dates[agent_acq_dates$AcquisitionDate > "2016-9-30 UTC",]
acq <- acq %>%
  filter(!is.na(AcquisitionDate))
summary(acq)
```


# Joining acquisition date with the new table

```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid <- leads_zuid %>%
  left_join(acq, c("ZUID" = "ZUID"))

```

```{r, warning=FALSE, message=FALSE}

#leads_zuid$LeadDateTime <- as.Date(leads_zuid$LeadDateTime)

```


# Too many searches for one RE Agent?

```{r, warning=FALSE, message=FALSE, echo=TRUE}
n_occur <- data.frame(table(leads_zuid$REAgentID))
n_occur[n_occur$Freq > 200,]

```


# Remove the ones with more than 200 unique leads and those without any REAgentID and ZUID

```{r, echo=TRUE}
leads_zuid <- leads_zuid %>%
  filter(ZUID != 1)

leads_zuid <- leads_zuid %>%
  filter(!REAgentID %in% c("0033300001ekvDDAAY", "0033300001tuTJeAAM", "0034000001JUSvpAAH", "0034000001SEjLlAAL"))

```


# Rename LeadType to Organic and Email

```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid$LeadVendor[leads_zuid$LeadType == "Social Organic"] <- "Organic"
leads_zuid$LeadVendor[leads_zuid$LeadType == "Organic"] <- "Organic"
leads_zuid$LeadVendor[leads_zuid$LeadType == "Email"] <- "Email"

```


```{r, warning=FALSE, message=FALSE, echo=TRUE}
skim(leads_zuid)

```


# Unique LeadType

```{r, warning=FALSE, message=FALSE, echo=TRUE}

unique(leads_zuid$LeadType)

```


# Unique LeadVendors

```{r, warning=FALSE, message=FALSE, echo=TRUE}

unique(leads_zuid$LeadVendor)

```



```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid %>%
  select(LeadVendor) %>%
  count(LeadVendor) %>%
  filter(n > 3)
  

```


# Grouping together into Leadtype

```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid <- leads_zuid %>%
  mutate(LeadType = case_when(
    LeadType %in% c("Social Organic", "Organic") ~ "Organic",
    LeadType == "Email" ~ "Email",
    LeadType == "Paid" ~ "Paid",
    LeadType == "unknown" ~ "Unknown",
    is.na(LeadType) ~ "NA",
    TRUE ~ "Unknown")
    )
```


# Group togethher lead channel with what we have cost for

```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid <- leads_zuid %>%
  mutate(LeadVendor = case_when(
    LeadVendor %in% c("email_welcome", "email", "Gmail", "email_autoresponder", "gmail",
                      "email_engagement", "Email", "email_content", "Yahoo", "Email_sales", "Email",
                      "email_sales_june_2015", "email_sales", "email_content_august_2014",
                      "email_sales_nov_2014", "email_sales_july_2014", "email_content_apr_2015", 
                  "email_content_july_2014", "hs_email", "Outlook", "email_content_feb_2014", "email_content_apr_2014",
                      "email_product") ~ "Email", 
    LeadVendor %in% c("Facebook", "facebook", "Linkedin Ads", "Instagram", "Linkedin Display", "instagram",
                      "Linkedin", "Twitter", "twitter") ~ "Social",
    LeadVendor == "Bing Search" ~ "Bing",
    LeadVendor == "Google Search" ~ "Google",
    LeadVendor %in% c("Affiliate", "zillow", "internal", "organic", "Zillow", "Trulia", "Trulia_R4R", "Organic") ~ "Organic",
    is.na(LeadVendor) ~ "NA",
    TRUE ~ "Other")
    )

```


```{r, warning=FALSE, message=FALSE, echo=TRUE}

unique(leads_zuid$LeadVendor)

```

# Count of lead Vendor

```{r, echo=TRUE}

sqldf('select LeadVendor, count(*) from leads_zuid group by 1')

```


# Marketing activity

# Leads can happen after acquisition date (roughly 30%), remove those rows.

```{r, echo=TRUE}

# Create a new variable LeadTimes
leads_zuid["x"]  <- 1
library(data.table)
DT <- data.table(leads_zuid)

DT[, LeadTimes:= cumsum(x), by = list(REAgentID)]

DT<- dplyr::select(DT, -c(x))
# calculate the average LeadTimes when people got acquisited
DTe <- DT %>% filter(!is.na(AcquisitionDate))
DTe <- DTe %>% mutate(dif = AcquisitionDate - LeadDateTime)

# 29% (3585 / 12341) acuiqistion happen before lead 
newdf <- subset(DTe, dif< 0) 
newdf <- subset(newdf, LeadTimes == 1 )
count(newdf) # 3585
length(unique(DTe$ZUID)) #12341

# only select acquistion happened after leads
DTe <- DTe %>% filter(dif >0)

DTe <- DTe %>% group_by(REAgentID) %>% slice(which.min(dif))
mean(DTe$LeadTimes) # 1.43
median(DTe$LeadTimes) # 1
summary(DTe)

nrow(DTe[DTe$LeadTimes == 1,]) # 6292
# 72% (6292 / 8756) of people got acquisited after the first lead

```


# Sales activity

```{r, warning=FALSE, message=FALSE, echo=TRUE}
phonecalls_16_17 <- phonecalls_16_17 %>%
  filter(TalkTimeMinutes != 0) %>%
  filter(TalkTimeMinutes < 60)

phonecalls_17 <- phonecalls_17 %>%
  filter(TalkTimeMinutes != 0) %>%
  filter(TalkTimeMinutes < 60)

```


```{r, warning=FALSE, message=FALSE, echo=TRUE}

summary(phonecalls_16_17)

```


# Inbound and Outbound phone calls

```{r, echo=TRUE}
phonecalls_16_17 <- phonecalls_16_17 %>%
  mutate(Activity = case_when(PhoneCallType == "RE Agent Called Sales Rep" ~ "Inbound",
                              PhoneCallType == "Sales Rep Called RE Agent" ~ "Outbound"))
	
phonecalls_17 <- phonecalls_17 %>%
 mutate(Activity = case_when(PhoneCallType == "RE Agent Called Sales Rep" ~ "Inbound",
                              PhoneCallType == "Sales Rep Called RE Agent" ~ "Outbound"))
```

```{r, warning=FALSE, message=FALSE, echo=TRUE}
phonecall <- rbind(phonecalls_16_17, phonecalls_17)

```

```{r, warning=FALSE, message=FALSE, echo=TRUE}
phonecall <- phonecall %>%
  select(-SalesRepID, -PhoneCallType)

```

# Remove NA rows in phonecall

```{r, echo=TRUE}
phonecall <- phonecall %>%
  filter(REAgentID != "")

```

# Join Sales Meeting and Phone calls just to monitor activity of REAgentID
# Since, Joining them will create duplicate sales meeting activity we have used phone calls and sales activity to calculate cost

```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid_phonecall_salesmeeting <- leads_zuid %>%
  left_join(phonecall, c("REAgentID" = "REAgentID"))

```

# Number of phone calls by Lead vendor

```{r, echo=TRUE}
sqldf('select LeadVendor, sum(TalkTimeMinutes), count(PhoneCallDateTime)
      from leads_zuid_phonecall_salesmeeting
      group by 1')

```


# Number of phone calls by Lead vendor - only for acquired customer

```{r, echo=TRUE}
sqldf('select LeadVendor, sum(TalkTimeMinutes), count(PhoneCallDateTime)
      from leads_zuid_phonecall_salesmeeting
      where AcquisitionDate is not NULL
      group by 1')
```


```{r, warning=FALSE, message=FALSE, echo=TRUE}
sales_meeting <- sales_meeting %>%
  select(-SalesRepID)

```

```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid_phonecall_salesmeeting <- leads_zuid_phonecall_salesmeeting %>%
  left_join(sales_meeting, c("REAgentID" = "REAgentID"))

```


```{r, echo=TRUE}

leads_zuid_phonecall_salesmeeting$PhoneCallDateTime <-  as.Date(leads_zuid_phonecall_salesmeeting$PhoneCallDateTime)

leads_zuid_phonecall_salesmeeting$SalesMeetingDate <- as.Date(leads_zuid_phonecall_salesmeeting$SalesMeetingDate)

```

# Select those where Acquisition Date >= Lead Date or Acquisition Date is null

```{r, echo=TRUE}
leads_zuid_final <- leads_zuid %>%
  filter(AcquisitionDate >= LeadDateTime | is.na(AcquisitionDate))

leads_zuid_phonecall_salesmeeting_final <- leads_zuid_phonecall_salesmeeting %>%
  filter(AcquisitionDate >= LeadDateTime | is.na(AcquisitionDate))

```


```{r, warning=FALSE, message=FALSE, echo=TRUE}

leads_zuid_phonecall_salesmeeting_final <- leads_zuid_phonecall_salesmeeting_final %>%
  filter(AcquisitionDate >= PhoneCallDateTime | is.na(AcquisitionDate))

```



# Meeting and Phone Calls Cost Assignment

```{r, warning=FALSE, message=FALSE, echo=TRUE}

leads_zuid_phonecall_salesmeeting_final <- sqldf('Select *, (SalesMeeting_Cost+SalesPhoneCall_Cost) as Sales_Cost
      from ( Select *,
      (case when SalesMeetingDate is not NULL then 300
      else 0 end) as SalesMeeting_Cost,
      case when PhoneCallDateTime is not NULL then 10
      else 0 end as SalesPhoneCall_Cost
      from leads_zuid_phonecall_salesmeeting_final)')

leads_zuid_phonecall_salesmeeting_final <- subset(leads_zuid_phonecall_salesmeeting_final, select = -c(SalesMeeting_Cost, SalesPhoneCall_Cost))

```


```{r, echo=TRUE}

sqldf('select LeadVendor, count(*) from leads_zuid_final group by 1')

```

# Graph of frequency by Lead Vendors

```{r, warning=FALSE, message=FALSE, echo=TRUE}
ggplot(data = leads_zuid) +
  aes(x = LeadVendor) +
  geom_bar(fill = "#0c4c8a") +
  theme_minimal()

leads_zuid %>%
  drop_na(AcquisitionDate) %>%
ggplot(aes(x = LeadVendor)) +
  geom_bar()

leads_zuid %>%
  mutate(AcquisitionDate = case_when(is.na(AcquisitionDate) ~ "NA",
                   TRUE ~ "Other")) %>%
  filter(AcquisitionDate == "NA") %>%
  group_by(LeadVendor) %>%
  ggplot(aes(x = LeadVendor)) +
  geom_bar()

```

# Graph of Lead Type vs LeadDateTime

```{r, warning=FALSE, message=FALSE, echo=TRUE}
ggplot(data = leads_zuid) +
  aes(x = LeadDateTime, fill = LeadType) +
  geom_histogram(bins = 30) +
  theme_minimal()
```



```{r, warning=FALSE, message=FALSE, echo=TRUE}
leads_zuid %>%
  mutate(AcquisitionDate = case_when(is.na(AcquisitionDate) ~ "NA",
                   TRUE ~ "Other")) %>%
  filter(AcquisitionDate == "NA") %>%
  group_by(LeadType) %>%
  ggplot(aes(x = LeadType)) +
  geom_bar()
  
```


# Find contribution of lead platform, lead type and lead vendor during the end to end conversation between Sales Agent and REAgentID


```{r, warning=FALSE, message=FALSE, echo=TRUE}

agent_zuid_leadPlatform_leadType_leadVendor <- sqldf('SELECT REAgentID, ZUID, 
      case when ZUID is NULL then "No" else "Yes" end as ZUID_Yes_No,
      case when AcquisitionDate is NULL then "NotAcquired" else "Acquired" end as AcqDate_Yes_No,
      sum(case when LeadPlatform = "Web" then 1 else 0 end) as count_LeadPlatform_web,  
      sum(case when LeadPlatform = "Phone" then 1 else 0 end) as count_LeadPlatform_phonecalls, 
      sum(case when LeadType = "Email" then 1 else 0 end) as count_LeadType_email, 
      sum(case when LeadType = "Paid" then 1 else 0 end) as count_LeadType_paid, 
      sum(case when LeadType = "Organic" then 1 else 0 end) as count_LeadType_organic, 
      sum(case when LeadType = "Unknown" then 1 else 0 end) as count_LeadType_unknown, 
      sum(case when LeadVendor = "Other" then 1 else 0 end) as count_LeadVendor_other, 
      sum(case when LeadVendor = "Email" then 1 else 0 end) as count_LeadVendor_email, 
      sum(case when LeadVendor = "Organic" then 1 else 0 end) as count_LeadVendor_organic, 
      sum(case when LeadVendor = "NA" then 1 else 0 end) as count_LeadVendor_na, 
      sum(case when LeadVendor = "Social" then 1 else 0 end) as count_LeadVendor_social, 
      sum(case when LeadVendor = "Google" then 1 else 0 end) as count_LeadVendor_google, 
      sum(case when LeadVendor = "Bing" then 1 else 0 end) as count_LeadVendor_bing,
      count(*) as number_of_meetings
      from leads_zuid_final
      group by 1, 2, 3')

percentage_leads_share <- sqldf('Select REAgentID, ZUID, ZUID_Yes_No,
(count_LeadPlatform_web * 1.0/ (count_LeadPlatform_web + count_LeadPlatform_phonecalls))*100 as percent_LeadPlatform_web,
(count_LeadPlatform_phonecalls * 1.0/ (count_LeadPlatform_web + count_LeadPlatform_phonecalls))*100 as percent_LeadPlatform_phonecalls,
(count_LeadType_email * 1.0/ (count_LeadType_email + count_LeadType_paid + count_LeadType_organic + count_LeadType_unknown))*100 as percent_LeadType_email,
(count_LeadType_paid * 1.0/ (count_LeadType_email + count_LeadType_paid + count_LeadType_organic + count_LeadType_unknown))*100 as percent_LeadType_paid,
(count_LeadType_organic * 1.0/ (count_LeadType_email + count_LeadType_paid + count_LeadType_organic + count_LeadType_unknown))*100 as percent_LeadType_organic,
(count_LeadType_unknown * 1.0/ (count_LeadType_email + count_LeadType_paid + count_LeadType_organic + count_LeadType_unknown))*100 as percent_LeadType_unknown,
(count_LeadVendor_other * 1.0/ (count_LeadVendor_other + count_LeadVendor_email + count_LeadVendor_organic + count_LeadVendor_na + count_LeadVendor_social + count_LeadVendor_google + count_LeadVendor_bing))*100 as percent_LeadVendor_other,
(count_LeadVendor_email * 1.0/ (count_LeadVendor_other + count_LeadVendor_email + count_LeadVendor_organic + count_LeadVendor_na + count_LeadVendor_social + count_LeadVendor_google + count_LeadVendor_bing))*100 as percent_LeadVendor_email,
(count_LeadVendor_organic * 1.0/ (count_LeadVendor_other + count_LeadVendor_email + count_LeadVendor_organic + count_LeadVendor_na + count_LeadVendor_social + count_LeadVendor_google + count_LeadVendor_bing))*100 as percent_LeadVendor_organic,
(count_LeadVendor_na * 1.0/ (count_LeadVendor_other + count_LeadVendor_email + count_LeadVendor_organic + count_LeadVendor_na + count_LeadVendor_social + count_LeadVendor_google + count_LeadVendor_bing))*100 as percent_LeadVendor_na,
(count_LeadVendor_social * 1.0/ (count_LeadVendor_other + count_LeadVendor_email + count_LeadVendor_organic + count_LeadVendor_na + count_LeadVendor_social + count_LeadVendor_google + count_LeadVendor_bing))*100 as percent_LeadVendor_social,
(count_LeadVendor_google * 1.0/ (count_LeadVendor_other + count_LeadVendor_email + count_LeadVendor_organic + count_LeadVendor_na + count_LeadVendor_social + count_LeadVendor_google + count_LeadVendor_bing))*100 as percent_LeadVendor_google,
(count_LeadVendor_bing * 1.0/ (count_LeadVendor_other + count_LeadVendor_email + count_LeadVendor_organic + count_LeadVendor_na + count_LeadVendor_social + count_LeadVendor_google + count_LeadVendor_bing))*100 as percent_LeadVendor_bing
from agent_zuid_leadPlatform_leadType_leadVendor group by 1, 2, 3')

percentage_leads_share[is.na(percentage_leads_share)] <- 0

phonecall_totalTalkTime <- sqldf('SELECT REAgentID, 
                                 sum(case when Activity = "Inbound" then 1 else 0 end) as Inbound,
                                 sum(case when Activity = "Outbound" then 1 else 0 end) as Outbound,                  
                                 sum(case when Activity = "Inbound" then TalkTimeMinutes else 0 end) as Inbound_total_talktime,
           sum(case when Activity = "Outbound" then TalkTimeMinutes else 0 end) as Outbound_total_talktime,
           sum(TalkTimeMinutes) from phonecall group by 1')



```


```{r, echo=TRUE}

agent_zuid_phonecall <- sqldf('Select * from agent_zuid_leadPlatform_leadType_leadVendor a left join phonecall_totalTalkTime b on a.REAgentID = b.REAgentID')

agent_zuid_phonecall[is.na(agent_zuid_phonecall)] <- 0

```




```{r, echo=TRUE}
#set the first row

agent_zuid_leadPlatform_leadType_leadVendor[1, "REAgentID"] <- 000
agent_zuid_leadPlatform_leadType_leadVendor[1, "ZUID"] <- 000
```



```{r, echo=TRUE}
write.csv(agent_zuid_phonecall, file = "Final.csv")
```

```{r, echo=TRUE}

meeting_complete <- agent_lookup %>%
  left_join(agent_acq_dates, c("ZUID" = "ZUID"))

```


```{r, message=FALSE, warning=FALSE, echo=TRUE}
meeting_complete <- meeting_complete %>%
  left_join(sales_meeting, c("REAgentID" = "REAgentID"))

```

```{r, echo=TRUE}

meeting_complete <- meeting_complete %>%
  drop_na(SalesMeetingDate)

```

```{r, echo=TRUE}
meeting_complete <- sqldf('SELECT REAgentID, ZUID, AcquisitionDate, SalesMeetingDate,
      (case when SalesMeetingDate is not NULL then 300
      else 0 end) as SalesMeeting_Cost
      from meeting_complete')


```

```{r, echo=TRUE}
meeting_complete <- sqldf('SELECT REAgentID, ZUID, 
      case when AcquisitionDate is NULL then "NotAcquired" else "Acquired" end as AcqDate_Yes_No,
      SalesMeeting_Cost
      from meeting_complete
      group by REAgentID')


```


```{r, echo=TRUE}

ggplot(data = meeting_complete) +
  aes(x = AcqDate_Yes_No) +
  geom_bar(fill = "#0c4c8a") +
  theme_minimal()

```

```{r, echo=TRUE}

agent_zuid_leadPlatform_leadType_leadVendor %>%
  select(REAgentID) %>%
  count(REAgentID) %>%
  filter(n > 1)

```


# most leads through the web
```{r, warning=FALSE, message=FALSE, echo=TRUE}
ggplot(data = agent_leads) +
  aes(x = LeadPlatform) +
  geom_bar(fill = "#0c4c8a") +
  theme_minimal()
```

# Organic gives the most leads
```{r, warning=FALSE, message=FALSE, echo=TRUE}
ggplot(data = agent_leads) +
  aes(x = LeadType) +
  geom_bar(fill = "#0c4c8a") +
  theme_minimal() +
  facet_wrap(vars(LeadPlatform))

```


# Organic clearly the best for acqusition as well

```{r, warning=FALSE, message=FALSE, echo=TRUE}

leads_zuid %>%
  drop_na(AcquisitionDate) %>%
  drop_na(LeadType) %>%
  group_by(LeadPlatform, LeadType) %>%
  ggplot(aes(x = LeadPlatform, fill = LeadType)) +
  geom_bar(position = "dodge")

```




```{r, echo=TRUE}

leads_zuid_phonecall_salesmeeting_final %>%
  filter(!is.na(TalkTimeMinutes)) %>%
  group_by(REAgentID) %>%
  ggplot(aes(TalkTimeMinutes, fill = Activity)) + 
  geom_density(alpha = 0.6)
  


```

# ROI calculation

```{r, echo=TRUE}
# for missing value in LeadType, assign average NPV
last_touch <- DTe %>%
  mutate(Revenue = ifelse(LeadType == "Paid", 7000,
                          ifelse(LeadType == "Email", 8000,
                                 ifelse(LeadType == "Organic", 7500, 7500))))

last_touch <- dplyr::select(last_touch, c(REAgentID, Revenue))

# Join last_touch table with percentage_leads_share

Revenue <- percentage_leads_share %>% left_join(last_touch, c("REAgentID" = "REAgentID"))

# only select paying customer rows
paying_customer <- Revenue %>% filter(!is.na(Revenue))
paying_customer$REAgentID<-as.factor(paying_customer$REAgentID)
summary(paying_customer)



paying_customer <- paying_customer %>% 
  mutate(EmailRevenue = Revenue * percent_LeadType_email/100) %>%
  mutate(PaidRevenue = Revenue * percent_LeadType_paid/100) %>%
  mutate(OrganicRevenue = Revenue * percent_LeadType_organic/100) %>%
  mutate(UnknownRevenue = Revenue * percent_LeadType_unknown /100) %>%
  mutate(Lead_Vendor_otherRevenue = Revenue * percent_LeadVendor_other/100) %>%
  mutate(Lead_Vendor_emailRevenue = Revenue * percent_LeadVendor_email/100) %>%
  mutate(Lead_Vendor_organicRevenue = Revenue * percent_LeadVendor_organic/100) %>%
  mutate(Lead_Vendor_naRevenue = Revenue * percent_LeadVendor_na/100) %>%
  mutate(Lead_Vendor_socialRevenue = Revenue * percent_LeadVendor_social/100) %>%
  mutate(Lead_Vendor_bingRevenue = Revenue * percent_LeadVendor_bing/100) %>%
  mutate(Lead_Vendor_googleRevenue = Revenue * percent_LeadVendor_google/100) 
  


paying_customer <- dplyr::select(paying_customer, c(REAgentID,Lead_Vendor_otherRevenue, Lead_Vendor_emailRevenue, Lead_Vendor_organicRevenue, Lead_Vendor_naRevenue, Lead_Vendor_socialRevenue, Lead_Vendor_bingRevenue, Lead_Vendor_googleRevenue,Revenue))
summary(paying_customer)

# marketing cost
agent_zuid_leadPlatform_leadType_leadVendor <- agent_zuid_leadPlatform_leadType_leadVendor %>%
  mutate(PaidSocialCost = 150 * count_LeadVendor_social) %>%
  mutate(PaidBingCost = 80 * count_LeadVendor_bing) %>%
  mutate(PaidGoogleCost = 100 * count_LeadVendor_google) %>%
  mutate(EmailCost = 10 * count_LeadVendor_email ) %>%
  mutate(OrganicCost = 0 * count_LeadType_organic ) %>%
  mutate(UnknownCost = 60 * count_LeadType_unknown) %>%
  mutate(NACost = 60 * count_LeadVendor_na)

# ROI - Marketing Channel

ROIpre <- agent_zuid_leadPlatform_leadType_leadVendor %>% left_join(paying_customer, c("REAgentID" = "REAgentID"))

SumSocialCost <- sum(ROIpre$PaidSocialCost)
SumBingCost <- sum(ROIpre$PaidBingCost)
SumGoogleCost <- sum(ROIpre$PaidGoogleCost)
SumEmailCost <- sum(ROIpre$EmailCost)
SumOrganicCost <- sum(ROIpre$OrganicCost)

SumSocialRevenue <- sum(ROIpre$Lead_Vendor_socialRevenue, na.rm=TRUE)
SumBingRevenue <- sum(ROIpre$Lead_Vendor_bingRevenue, na.rm=TRUE)
SumGoogleRevenue <- sum(ROIpre$Lead_Vendor_googleRevenue, na.rm=TRUE)
SumEmailRevenue <- sum(ROIpre$Lead_Vendor_emailRevenue, na.rm=TRUE)
SumOrganicRevenue <- sum(ROIpre$Lead_Vendor_organicRevenue, na.rm=TRUE)

ROISocial = (SumSocialRevenue - SumSocialCost) / SumSocialCost
ROIBing = (SumBingRevenue - SumBingCost)/SumBingCost
ROIGoogle = (SumGoogleRevenue - SumGoogleCost)/SumGoogleCost
ROIEmail = (SumEmailRevenue - SumEmailCost)/SumEmailCost
ROIOrganic = (SumOrganicRevenue - SumOrganicCost)/SumOrganicCost

              
```



```{r, echo=TRUE}
# ROI Sales Channel
sales <- dplyr::select(leads_zuid_phonecall_salesmeeting_final, c(REAgentID, Activity))
summary(sales)
```


```{r, echo=TRUE}
#phone call cost per REAgentID
sales["phonecallcost"] <- 10

sales <- sales%>% group_by(REAgentID) %>%
  summarize(SumPhoneCallCost = sum(phonecallcost))

#meeting cost per REAgentID
sales_meeting["meetingcost"] <- 300
sales_meeting <- sales_meeting %>% group_by(REAgentID) %>%
  summarize(SumMeetingCost = sum(meetingcost))

# Join table
Sales_final <- sales %>% left_join(sales_meeting, c = ("REAgentID" = "REAgentID"))

ROIpre <- dplyr::select(ROIpre, c(REAgentID, AcqDate_Yes_No,Revenue))
Sales_final <- ROIpre %>% left_join(Sales_final,c = ("REAgentID" = "REAgentID") )

Sales_final <- Sales_final[!with(Sales_final,is.na(SumMeetingCost)&is.na(SumPhoneCallCost)),]
sales_pre <- Sales_final %>% filter(!is.na(Revenue))%>% 
  mutate(SumPhoneCallCost = ifelse(is.na(SumPhoneCallCost),0,SumPhoneCallCost)) %>%
  mutate(SumMeetingCost = ifelse(is.na(SumMeetingCost),0,SumMeetingCost)) 
sales_pre <- sales_pre %>%
  mutate(PhoneCallRevenue = Revenue * (SumPhoneCallCost/(SumPhoneCallCost+SumMeetingCost))) %>%
           mutate(MeetingRevenue = Revenue * (SumMeetingCost/(SumPhoneCallCost+SumMeetingCost)))

Sales_final <- Sales_final %>% left_join(sales_pre, c = ("REAgentID" = "REAgentID"))

# Sales Chanel ROI

SumPhoneCallCost <- sum(Sales_final$SumPhoneCallCost, na.rm=TRUE)
SumMeetingCost <- sum(Sales_final$SumMeetingCost, na.rm=TRUE)
SumPhoneCallRevenue <- sum(Sales_final$PhoneCallRevenue, na.rm=TRUE)
SumMeetingRevenue <- sum(Sales_final$MeetingRevenue, na.rm=TRUE)

ROIPhoneCall <- (SumPhoneCallRevenue - SumPhoneCallCost)/SumPhoneCallCost
ROIMeeting <- (SumMeetingRevenue - SumMeetingCost) / SumMeetingCost
```

