---
title: "Findings related to House Sales in King County"
author: "By Nancy Jain"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r echo = FALSE}

# Title: Presentation
# Purpose: Findings related to effectiveness of Advertising and Promotions
# Date: March 3, 2019
# Author:Nancy Jain
```

```{r echo = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

```{r echo = FALSE, warning=FALSE, message=FALSE}

# load libraries and plots
library(plotly)
library(stringr)


load("map_price_area.RData")
load("house_sales_by_date.RData")
load("sqft_living_lot_condition.RData")
load("zipcode_avg_price_per_sqft.RData")

```


Introduction  
=========================================
+ King County and its present
+ Expectation vs reality of King County housing market
+ Data available for King County housing market analysis study


Even though number of people migrating to King County have increased, house sales have decreased
=========================================

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplotly(house_sales_by_date)

```


Finding 1: Most of the houses sold have average condition, that is, condition = "3"
=========================================

```{r echo = FALSE, include = TRUE, warning=FALSE}

# Show the plot that was made in the Technical Appendix and saved in the workspace
sqft_living_lot_condition

```


Finding 2: Next, I have categorized the areas of King County relative to the house prices in that area. For a given price range, the graph shows all the areas with houses available in it.
=========================================


```{r echo = FALSE, include = TRUE, warning=FALSE}

map_price_area

```


================================================

```{r echo = FALSE, include = TRUE, warning=FALSE}

zipcode_avg_price_per_sqft

```

Conclusion
===============

+ Top takeaways
    - What have we missed and what should we focus?
    - What is changing which we should accomodate in decision making?

## Thank You

